---
title: "Course ML - Project"
author: "Klod Kolaro"
date: "November 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions:
exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), 
lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).
Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes.
The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience.  
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

```{r }
options(warn=-1)
library(caTools)
library(caret)
library(ggplot2)
library(dplyr)
library(sqldf)
```


# Data import


```{r }
getwd()
setwd("C:/Users/kkolaro/Desktop/Private/ucenje/Coursera/machine learning/project")
data_training<-read.csv("pml-training.csv",na.strings = c("NA","#DIV/0!", " "))
dim(data_training)
data_test<-read.csv("pml-testing.csv",na.strings = c("NA","#DIV/0!", " "))
set.seed(12345)
```


#Exploratory  analysis


```{r }
unique(data_training$classe)
unique(data_training$user_name)
summary(data_training$classe)

# Total number of exercises per classe

sqldf("select count(classe) as Total_number,classe from data_training group by classe order by Total_number desc")
summary(data_training[,c("classe")])

barplot(table(data_training$classe), main = "Classe distribution", beside=FALSE,xlab="Clasee",ylab="Number of activity",col="red")

# Total number of exercises  per user

sqldf("select count(classe) as Total_number,user_name from data_training group by user_name order by Total_number desc")

# Participants per classes

sqldf("select count(classe),user_name,classe from data_training group by user_name,classe")
usr_class<-prop.table(table(data_training$classe,data_training$user_name),2)#procentage of user perclass
print(usr_class)

barplot(usr_class, main="Distribution of participants activities per clasee",xlab = "User",legend = row.names(usr_class),beside=TRUE,col=c("darkblue","red","green","cyan","white"),,xlim=c(0, ncol(usr_class) + 37))

# Classes per participants

class_usr<-prop.table(table(data_training$user_name,data_training$classe),2)#class per user
print(class_usr)
barplot(class_usr, main="Classe distribution per paricipants",xlab = "Classe",legend = row.names(class_usr),beside=TRUE,col=c("darkblue","red","green","cyan","white","yellow"),xlim=c(0, ncol(class_usr) + 39))
```

## Key findings:

Max number of activities is of classe A - 5580 and min of classe D - 3216
Out of total number of A classe ,Jeremy contribute with  21,1%  ,and  Pedro only 11,4%
The highest % of class A  had Jeremy 34,59% and the lowest Pedro 24,5%


#Data cleansing & transformation


```{r }
data_training_NA<-sapply(data_training, function(x)all(!is.na(x)))
data_test_NA<-sapply(data_test, function(x)all(!is.na(x)))

data_training<-data_training[,data_training_NA]
data_test<-data_test[,data_test_NA]

data_training<-subset(data_training, select = -c(X,user_name))
data_test<-subset(data_test, select = -c(X,user_name))

data_training<-data_training[,!(grepl("window|timestamp", colnames(data_training)))]
data_test<-data_test[,!(grepl("window|timestamp", colnames(data_test)))]
```


# Modeling


```{r }

# Creating test and training datasets from data_training set and  Cross validation

inTrain<-createDataPartition(data_training$classe,p=0.7,list = FALSE)
training<-data_training[inTrain,]
testing<-data_training[-inTrain,]

# RF-willuse a Random forests a method well-suited to problems with a relatively small number of 
# observations with a large number of feature variables.

controlPar<-trainControl(method = "cv",number=5)
Random_model<-train(classe~.,data=training,method="rf", trControl=controlPar,ntree=50)

Random_model

prediction<-predict(Random_model,testing)
table(predictions=prediction, actual=testing$classe)
```


# Predicting  20 different test cases


```{r }
prediction_test<-predict(Random_model,data_test)
print(prediction_test)

options(warn=0)
```
